Cassini State Capture

### 2019-08-27 writeups/Su_sep_cross.tex
typo fix
fix sign mistake, problem 3 looks good

We then seek $\Delta_{\pm}$, determined by
\begin{align}
    \Delta_{\pm} &= \int\limits_{\mathcal{C}_{\pm}}
            \pd{h}{\eta}\eta'\;\mathrm{d}\tau,\nonumber\\
        &= \int\limits_{\mathcal{C}_{\pm}}
            \frac{\pd{h}{\eta}}{\phi'}\eta'\;\mathrm{d}\phi.
\end{align}
We must proceed carefully since $\phi' \to 0$ at the endpoints of
$\mathcal{C}_{\pm}$; careful work shows these zeros cancel against zeros of
$\pd{h}{\eta}$, and
\begin{align}
    \Delta_{\pm} &= \int\limits_{\mathcal{C}_{\pm}} \eta'
        \p{\cos I \mp \sin I \sqrt{\frac{1 - \cos \phi}{2\eta\sin I}}}
            \;\mathrm{d}\phi,\label{eq:p1_delta_eq}\\
        &= \mp 2\pi \epsilon \eta \cos I
            + 4\epsilon\sqrt{\eta \sin I}.\label{eq:p1_delta}
\end{align}
Call $\eta_\star$ the value of $\eta$ at which a trajectory encounters the
separatrix. There are then three possibilities:
\begin{itemize}
    \item For $\eta_\star < \frac{4\sin I}{\pi^2 \cos^2 I}$, we obtain
        $\Delta_{\pm} > 0$. In this regime, all separatrix encounters
        originating in zones I/III result in secure capture (see
        \autoref{ss:z_out_ic}), while all trajectories originating in zone II do
        not experience separatrix encounter. For $I = 20^\circ$, this condition
        equates to $\eta_\star \lesssim 0.16$. This condition is not to be taken
        too exactly, as all our calculations are performed in the small $\eta$
        limit. Select simulations in $\eta_\star \lesssim 0.1$ are in agreement
        with the secure capture prediction. An example of such a simulation is
        presented in \autoref{fig:2_1testo4}.
        \begin{figure}[t]
            \centering
            \includegraphics[width=\columnwidth]{../initial/2_toy2/1testo4.png}
            \caption{Simulation with initial condition in zone I subject to
            $\eta' = \epsilon \eta$, where $\epsilon = 10^{-4}$. The top panel
            shows the trajectory (red) starting with initial condition $\mu_0 =
            0.140, \phi_0 = 0$ (green dot) and separatrix at separatrix crossing
            (dashed black) in the labelled coordinates. The end of the
            simulation is labelled in the blue cyan dot. The bottom panel shows
            the evolution of the enclosed area by the trajectory (red dots),
            area of the separatrix (blue) and $\eta$ (dotted green) over
            time.}\label{fig:2_1testo4}
        \end{figure}

    \item If $\eta_\star > \frac{4\sin I}{\pi^2 \cos^2 I}, \eta_\star < \eta_c$,
        then the separatrix still exists, and $\Delta_+ < 0, \Delta_- > 0$,
        $\Delta_+ + \Delta_- > 0$. This implies: zone I + zone II trajectories
        do not experience separatrix encounters, while zone III trajectories are
        captured into zone II with probability
        \begin{equation}
            P_c = \frac{8\sqrt{\eta_\star \sin I}}{2\pi \eta_\star \cos I +
                4\sqrt{\eta_\star \sin I}}.
        \end{equation}

@@

If instead $\epsilon < 0$, then the separatrix area shrinks over time. For
initial values $\eta > \eta_c$, the separatrix appears over time and the
scenario is analyzed in \citep{anderson2018teeter}.

If $\eta$ decreases past critical value $\frac{4\sin I}{\pi^2 \cos^2 I}$, i.e.\
$\Delta_{\pm} < 0$, then the separatrix appears. Here, trajectories in zones
I/III do not experience separatrix encounter while trajectories in zone II
transition into zones I/III with probabilities given respectively (see
\autoref{ss:z2_ic}):
\begin{equation}
    \begin{split}
        P_{II \to I} &= \frac{2\pi \eta_\star \cos I + 4\sqrt{\eta_\star \sin I}}
            {8\sqrt{\eta_\star \sin I}}, \\
        P_{II \to III} &= \frac{-2\pi \eta_\star \cos I + 4\sqrt{\eta_\star \sin
            I}} {8\sqrt{\eta_\star \sin I}}.
    \end{split}
\end{equation}

@@

We now consider the combined problem, where the EOMs for Problem 2
(\autoref{se:p2_eom}) are used in conjunction with the simplest $\eta' = \delta
\eta$, the same as \autoref{eq:p1_eom}. Evaluation of $\Delta h$ along
$\mathcal{C}_{\pm}$ is simply combining the results of \autoref{eq:p1_delta} and
\autoref{eq:p2_delta}. The result is
\begin{align}
    \Delta_{\pm} ={}& \epsilon\s{
        \mp 2\pi \p{1 - 2\eta \sin I - \eta^2 \cos^2 I}
            + 16\cos I\eta^{3/2}\sqrt{\sin I}}\nonumber\\
        &+ \delta \s{\mp 2\pi \eta \cos I + 4\sqrt{\eta \sin I}}.
\end{align}
Thus, the capture probability from a zone III initial condition into the
separatrix is
\begin{equation}
    P_c \approx \frac{\p{32 \eta_\star \cos I + 8f}\sqrt{\eta_\star \sin I}}{
        2\pi\p{1 - 2\eta_{\star}\sin I + f\eta_\star \cos I}
            + \p{16\eta_\star\cos I + 4f} \sqrt{\eta_\star\sin I}}.
            \label{eq:p3_p_c}
\end{equation}
An $\eta_\star^2$ term has been dropped from the denominator, and $f \equiv
\frac{\delta}{\epsilon}$ measures the relative strength of the two
perturbations. Note that $f = 0$ reduces to \autoref{s:p2}.

@@

\bibliographystyle{mnras}
\bibliography{paper}

>>>>

We then seek $\Delta_{\pm}$, determined by
\begin{align}
    \Delta_{\pm} &= \int\limits_{\mathcal{C}_{\pm}}
            \pd{h}{\eta}\eta'\;\mathrm{d}\tau,\nonumber\\
        &= \int\limits_{\mathcal{C}_{\pm}}
            \frac{\pd{h}{\eta}}{\phi'}\eta'\;\mathrm{d}\phi.
\end{align}
We must proceed carefully since $\phi' \to 0$ at the endpoints of
$\mathcal{C}_{\pm}$; careful work shows these zeros cancel against zeros of
$\pd{h}{\eta}$, and
\begin{align}
    \Delta_{\pm} &= \int\limits_{\mathcal{C}_{\pm}} \eta'
        \p{-\cos I \mp \sin I \sqrt{\frac{1 - \cos \phi}{2\eta\sin I}}}
            \;\mathrm{d}\phi,\label{eq:p1_delta_eq}\\
        &= \mp 2\pi \epsilon \eta \cos I
            + 4\epsilon\sqrt{\eta \sin I}.\label{eq:p1_delta}
\end{align}
Call $\eta_\star$ the value of $\eta$ at which a trajectory encounters the
separatrix. There are then three possibilities:
\begin{itemize}
    \item For $\eta_\star < \frac{4\sin I}{\pi^2 \cos^2 I}$, we obtain
        $\Delta_{\pm} > 0$. In this regime, all separatrix encounters
        originating in zones I/III result in secure capture (see
        \autoref{ss:z_out_ic}), while all trajectories originating in zone II do
        not experience separatrix encounter. For $I = 20^\circ$, this condition
        equates to $\eta_\star \lesssim 0.16$. This condition is not to be taken
        too exactly, as all our calculations are performed in the small $\eta$
        limit. Select simulations in $\eta_\star \lesssim 0.1$ are in agreement
        with the secure capture prediction. An example of such a simulation is
        presented in \autoref{fig:2_1testo4}.
        \begin{figure}[t]
            \centering
            \includegraphics[width=\columnwidth]{../initial/2_toy2/1testo4.png}
            \caption{Simulation with initial condition in zone I subject to
            $\eta' = \epsilon \eta$, where $\epsilon = 10^{-4}$. The top panel
            shows the trajectory (red) starting with initial condition $\mu_0 =
            0.140, \phi_0 = 0$ (green dot) and separatrix at separatrix crossing
            (dashed black) in the labelled coordinates. The end of the
            simulation is labelled in the blue cyan dot. The bottom panel shows
            the evolution of the enclosed area by the trajectory (red dots),
            area of the separatrix (blue) and $\eta$ (dotted green) over
            time.}\label{fig:2_1testo4}
        \end{figure}

    \item If $\eta_\star > \frac{4\sin I}{\pi^2 \cos^2 I}, \eta_\star < \eta_c$,
        then the separatrix still exists, and $\Delta_- < 0, \Delta_+ > 0$,
        $\Delta_+ + \Delta_- > 0$. This implies: zone III + zone II trajectories
        do not experience separatrix encounters, while zone I trajectories are
        captured into zone II with probability
        \begin{equation}
            P_c = \frac{8\sqrt{\eta_\star \sin I}}{2\pi \eta_\star \cos I +
                4\sqrt{\eta_\star \sin I}}.
        \end{equation}

@@

If instead $\epsilon < 0$, then the separatrix area shrinks over time. For
initial values $\eta > \eta_c$, the separatrix is not initially present and only
appears later in the simulation. This is analyzed in \citep{anderson2018teeter}.

If $\eta$ decreases past critical value $\frac{4\sin I}{\pi^2 \cos^2 I}$, i.e.\
$\Delta_{\pm} < 0$, then the separatrix appears. Here, trajectories in zones
I/III do not experience separatrix encounter while trajectories in zone II
transition into zones I/III with probabilities given respectively (see
\autoref{ss:z2_ic}):
\begin{equation}
    \begin{split}
        P_{II \to I} &= \frac{-2\pi \eta_\star \cos I + 4\sqrt{\eta_\star \sin I}}
            {8\sqrt{\eta_\star \sin I}}, \\
        P_{II \to III} &= \frac{2\pi \eta_\star \cos I + 4\sqrt{\eta_\star \sin
            I}} {8\sqrt{\eta_\star \sin I}}.
    \end{split}
\end{equation}

@@

We now consider the combined problem, where the EOMs for Problem 2
(\autoref{se:p2_eom}) are used in conjunction with the simplest $\eta' = \delta
\eta$, the same as \autoref{eq:p1_eom}. Evaluation of $\Delta h$ along
$\mathcal{C}_{\pm}$ is simply combining the results of \autoref{eq:p1_delta} and
\autoref{eq:p2_delta}. The result is
\begin{align}
    \Delta_{\pm} ={}& \epsilon\s{
        \mp 2\pi \p{1 - 2\eta \sin I - \eta^2 \cos^2 I}
            + 16\cos I\eta^{3/2}\sqrt{\sin I}}\nonumber\\
        &+ \delta \s{\pm 2\pi \eta \cos I + 4\sqrt{\eta \sin I}}.
\end{align}
Thus, the capture probability from a zone III initial condition into the
separatrix is
\begin{equation}
    P_c \approx \frac{\p{32 \eta_\star \cos I + 8f}\sqrt{\eta_\star \sin I}}{
        2\pi\p{1 - 2\eta_{\star}\sin I - f\eta_\star \cos I}
            + \p{16\eta_\star\cos I + 4f} \sqrt{\eta_\star\sin I}}.
            \label{eq:p3_p_c}
\end{equation}
An $\eta_\star^2$ term has been dropped from the denominator, and $f \equiv
\frac{\delta}{\epsilon}$ measures the relative strength of the two
perturbations. Note that $f = 0$ reduces to \autoref{s:p2}.

@@

\bibliographystyle{mnras}
\bibliography{Su_sep_cross}
